---
title: "Homework 2"
author: "Henry Stoddard"
date: "9/24/2020"
output: github_document
---
## Problem 1

Importing Mr Trash Wheel dataset and tidying 

```{r library, import, tidy}
library(tidyverse)
library(readxl)
trash_df = read_excel("./data/Mr_Trash_Wheel.xlsx")
trash_df = janitor::clean_names(trash_df) %>% select(-notes, -x16, -x17) %>% drop_na(dumpster) %>% mutate(sports_balls = round(sports_balls), sports_balls = as.integer(sports_balls))
```

Importing and tidying precipitation datasets
```{r precipitation}
precip2018_df = read_excel("./data/Mr_Trash_Wheel.xlsx", sheet = "2018 Precipitation", skip = 1) %>% janitor::clean_names() %>% drop_na(month) %>% mutate(year = 2018) %>% relocate(year)

precip2017_df = read_excel("./data/Mr_Trash_Wheel.xlsx", sheet = "2017 Precipitation", skip = 1) %>% janitor::clean_names() %>% drop_na(month) %>% mutate(year = 2017) %>% relocate(year)
```

Now combining datasets
```{r merge}
precip_df = bind_rows(precip2018_df, precip2017_df) %>% mutate(month = month.name[month])
```
```{r specific analysis}
sum(pull(precip2018_df, total)) %>% view()
balls_df = select(trash_df, sports_balls, year)
balls2017_df = subset(balls_df, year == 2017) %>%summarise_if(is.numeric, na.rm = TRUE, median) %>% view()
```

## Data Summary
These two datasets include info from the Mr. Trashwheel collector in Baltimore, MD. This machine collects trash and then stores it in a dumpster. Data includes info on year, month, trash collected, and the amounts of some specific kinds of trash. There are a total of `r nrow(trash_df)` rows in our final dataset. Additional data sheets include precipitation data by month and year, which have a total of `r nrow(precip_df)` rows. In 2018, there was a total of `r sum(pull(precip2018_df, total))` inches of rain. In 2017, the median number of sports balls in a dumpster was 8.